---
title: "test_mapping"
format: html
editor: visual
---


# Description

This qmd was used to map the study sites in the Bolinao-Anda Reef Complex, PH

# library

```{r}
#| label: libraries
#| output: false
#| eval: true
#| warning: false
#| message: false
#| cache: false

library(sf)
library(rio)
library(maps)
library(mapproj)
library(magrittr)
library(tidyverse) 
library(patchwork)
library(ggrepel)
library(ggspatial)
```

# base map for the PH

```{r}
ph <- map_data("world") |> 
  filter(region == "Philippines")
```

## plot basemap

```{r}
ph_map <-
ph |> 
  ggplot() +
  geom_polygon(aes(x = long,
                   y = lat,
                   group = group),
               fill = "grey40",
               color = "black",
               size = 0.3) +
  geom_rect(xmin = 119.3, 
            xmax = 120.5,
            ymin = 15.5,
            ymax = 17,
            color = "red",
            fill = NA, 
            linewidth = 1) +
  theme_classic() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 1))
  #coord_sf(ylim = c(16, 17),
   #        xlim = c(119, 121))
```

## Zoom into Bolinao

Lat and Long details are provided in the "ph" database

```{r}
bolinao_map <- 
ph |> 
  ggplot() +
  geom_polygon(aes(x = long,
                   y = lat,
                   group = group),
               fill = "grey40",
               color = "black",
               size = 0.3) +
  coord_sf(ylim = c(16, 16.8),
           xlim = c(119.7, 120.8)) +
  theme_classic() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank())
```

# load .shp file

Load shp files of municipalities in the Philippines Filter Bolinao

```{r}
shp_bolinao <- read_sf("Shape_files/MuniCities.shp") |> 
  filter(NAME_2 == "Bolinao")
```

## superimpose shp file on base map

```{r}
bolinao_shp_map <-
bolinao_map +
  geom_sf(data = shp_bolinao, fill = "gray40", color = "black", size = 0.1) +
  coord_sf(ylim = c(16.35, 16.48),
           xlim = c(119.8, 120.02)) +
  theme_classic() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 15),
        axis.line = element_blank()) +
  scale_y_continuous(name = "") +
  scale_x_continuous(name = "")
```

# inset PH map

```{r}
map_v1 <- 
bolinao_shp_map +
  annotation_custom(ggplotGrob(ph_map),
                    xmin = 119.97,
                    xmax = 120.02,
                    ymin = 16.42,
                    ymax = 16.48) +
  theme(panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 1))
```

# insert sites for BARC

## csv for sites

```{r}
bol_sites <- read_csv("Data/Sites_Lat_Long.csv",
                      trim_ws = TRUE)
```

## split El Pescador to manipulate its shape separately
```{r}
bol_assay_sites <-
  bol_sites |> 
  filter(Abbreviation != "ELP")

bol_collection_sites <-
  bol_sites |> 
  filter(Abbreviation == "ELP")
```


# superimpose BARC sites on base map

```{r}
map_v2 <- 
map_v1 +
  geom_point(data = bol_assay_sites,
             mapping = aes(x = Long,
                           y = Lat),
             pch = 21,
             color = "black",
             fill = "red2",
             size = 5) +
  geom_point(data = bol_collection_sites,
             mapping = aes(x = Long,
                           y = Lat),
             pch = 24,
             color = "black",
             fill = "blue",
             size = 5)
  #geom_text_repel(data = bol_sites,
   #               mapping = aes(x = Long,
    #                            y = Lat,
     #                           label = Site))
```

# add site name texts

```{r}
map_v3 <-
map_v2 +
  geom_text_repel(data = bol_assay_sites,
                  mapping = aes(x = Long,
                                y = Lat,
                                label = Site),
                  nudge_y = 0.01,
                  size = 3) +
  geom_text_repel(data = bol_collection_sites,
                  mapping = aes(x = Long,
                                y = Lat,
                                label = Site),
                  nudge_y = 0.01,
                  size = 3)
```

# add scale bars and arrows

```{r}
map_v4 <-
map_v3 +
  annotation_scale(location = "br",
                   width_hint = 0.2,
                   text_cex = 0.95) +
  annotation_north_arrow(location = "tl",
                         which_north = "true",
                         pad_x = unit(0.2, "in"),
                         pad_y = unit(0.2, "in"),
                         height = unit(0.9, "in"),
                         width = unit(0.9, "in"),
                         style = north_arrow_fancy_orienteering)
```

# add labels to Bolinao-Anda land

```{r}
map_v5 <-
map_v4 +
  annotate("text", x = 119.89, y = 16.37, label = "Bolinao", size = 6, color = "white") +
  annotate("text", x = 119.933, y = 16.40, label = "Santiago", size = 6, color = "white")
  
```

# ggsave

```{r}
ggsave("Plot/Fig. X. bolinao_map.jpeg",
       plot = map_v5, 
       dpi = 600, 
       height = 8, 
       width = 8, 
       units = "in")
```
