---
title: "Abundance_results_plots"
format: html
editor: visual
---

# Description

This qmd examines the abundance model fitted in the "fits" qmd file, and then visualizing the results.Also, runs code for generating a heatmap for each browser species across depth bands.

# MAXN ABUNDANCE

## load model

```{r}
load("Model_Outputs/abun_mod.Rdata")
```

## summary table

```{r}
abun_mod |> as_draws_df() |> 
  dplyr::select(matches("^b_.*")) |> 
  exp() |> 
  summarise_draws(median,
                  HDInterval::hdi,
                  Pl= ~mean(.x < 1),
                  Pg = ~mean(.x > 1)) |> 
  filter(Pl > 0.85 | 
           Pg > 0.85)
```

## depth bands newdata for plotting

```{r}
abun_nd <- 
  abun_mod |> 
  emmeans(~ Depth_band, 
          type = "link") |> 
  gather_emmeans_draws() |> 
  mutate(Exp_abun = exp(.value)) |> 
  dplyr::select(-.chain,
                -.iteration,
                -.value) |> 
  summarise_draws(median, 
                  HDInterval::hdi) |> 
  as.data.frame()
```

## plot depth band

```{r}
p_brow_abun <- 
abun_nd |> 
  ggplot(aes(x = Depth_band,
             y = median)) +
  geom_pointrange(aes(ymin = lower,
                      ymax = upper),
                  shape = 21,
                  size = 1.5,
                  fatten = 3,
                  color = "black",
                  fill = "red2") +
  scale_y_continuous(name = "Abundance (MaxN)",
                     limits = c(0, 2),
                     breaks = seq(0, 2, by = 1)) +
  scale_x_discrete(name = "") +
  theme_classic() +
    #ggtitle("Total herbivores") +
     theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y = element_text(size = 15),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 18)) +
  geom_image(data = tibble(Depth_band = 4,
                           median = 1.8),
             aes(image = "Images/naso-unicornis-bluespine-unicornfish.png"),
             size = 0.30) +
  geom_image(data = tibble(Depth_band = 4,
                           median = 1.4),
             aes(image = "Images/siganus-spp-rabbitfish.png"),
             size = 0.30)
```

## pairwise

```{r}
abun_mod |> 
emmeans(~ Depth_band,
        type = "link") |>
  pairs(reverse = TRUE) |> 
  gather_emmeans_draws() |> # odds-ratio
  mutate(.value = exp(.value)) |> # log odds-ratio  
  summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1))#,
            #.width = c(0.85, 0.95))
```

# HEATMAP

# summarize abundance per species and plot a heatmap

```{r}
heatmap_abun <-
brow_abun_v2 |> 
   mutate(Species = fct_reorder(Species,
         desc(Species))) |> 
  group_by(Depth_band, Species) |> 
  summarise(sumMaxN = sum(MaxN)) |> 
  filter(Species != "None") |> 
  mutate(Depth_band = recode(Depth_band,
                             "10m" = "0-10m",
                             "20m" = "11-20m",
                             "30m" = "31-40m",
                             "50m" = "41-50m")) |> 
  ggplot(aes(y = Species, x = Depth_band, fill = sumMaxN)) +
  geom_tile(color = "black",
            lwd = 0.75,
            linetype = 1) +
  scale_fill_gradient(low = "#DDD487",
                      high = "#E29297") +
  theme_classic() +
  theme(legend.position = "none") +
  ylab("MaxN relative abundance") +
  xlab("Depth bands") +
  theme(axis.text.y = element_text(face = "italic"))
```

# summary table of abundance per depth

```{r}
brow_abun_v2 |> 
   mutate(Species = fct_reorder(Species,
         desc(Species))) |> 
  group_by(Depth_band, Species) |> 
  summarise(mean = mean(MaxN))
```

## ggsave

```{r}
ggsave("Plots/Fig. X2. Heatmap.jpeg",
       plot = heatmap_abun, 
       dpi = 600, 
       height = 6,
       width = 6,
       units = "in")
```

### ends
