---
title: "sarg_assays_results_plots"
format: html
editor: visual
---

# Description

This qmd files examines the models from the "fit" qmd and visualizes the results.

# 4 DAY EXPERIMENT (96 h)

## load models

```{r}
load("Model_Outputs/logm_sarg96_mod.Rdata")
load("Model_Outputs/logm_sarg3h_mod.Rdata")
```

## FULL DATASET

## summary table

```{r}
logm_sarg96_mod |> as_draws_df() |> 
  dplyr::select(matches("^b_.*")) |> 
  exp() |> 
  summarise_draws(median,
                  HDInterval::hdi,
                  Pl= ~mean(.x < 1),
                  Pg = ~mean(.x > 1)) #|> 
  #filter(Pl > 0.85 | 
   #        Pg > 0.85)
```

## newdata for plotting macroalgae loss with depth band

```{r}
sarg_assay_96_nd <-
  logm_sarg96_mod |> 
  emmeans(~ Depth_band *
            Treatment, 
          type = "link") |> 
  gather_emmeans_draws() |> 
  mutate(Exp_grams = exp(.value)) |> 
  dplyr::select(-.chain,
                -.iteration,
                -.value) |> 
  summarise_draws(median, 
                  HDInterval::hdi) |> 
  as.data.frame()
```

## change depth band names

```{r}
sarg_assay_96_nd2 <-
sarg_assay_96_nd |> 
  mutate(Depth_band = recode(Depth_band,
                             "10m" = "0-10m",
                             "20m" = "11-20m",
                             "30m" = "31-40m",
                             "50m" = "41-50m"))

replicates <- c(32, 35, 35, 40, 29, 34, 7, 6)

sarg_assay_96_nd3 <- cbind(sarg_assay_96_nd2, 
                           replicates) |> 
  as.data.frame()

dodge_width <- 0.30
```

## plot 96h macroalgae loss with depth bands

```{r}
p_sarg_assay_96 <- 
sarg_assay_96_nd3 |> 
  ggplot(aes(x = Depth_band,
             y = median,
             fill = Treatment)) +
  geom_pointrange(aes(ymin = lower, 
                      ymax = upper,
                      shape = Treatment),
                  position = position_dodge(0.25),
                  size = 1.5,
                  fatten = 3) +
  scale_y_continuous(expression(italic(Sargassum)~loss~(grams~4~d^-1)),
                     limits = c(0,50),
                     breaks = seq(0, 50, by = 10)) +
  scale_x_discrete(name = "Depth bands") +
  theme_classic() +
  #ggtitle("Deployment duration: 96h") +
  scale_fill_manual(values = c("red2", "blue")) +
  scale_shape_manual(values = c(22, 21)) +
  theme(axis.title = element_text(size = 15),
        axis.text = element_text(size = 12),
        plot.title = element_text(size = 18)) +
  geom_image(data = tibble(Depth_band = 4.3,
                           median = 45,
                           Treatment = "Caged"),
             aes(image = "Images/sargassum-spp.png"),
             size = 0.16) +
    geom_text(aes(label = replicates,
                  y = upper + 1.5,
                  x = ),
              position = position_dodge(width = dodge_width),
              color = "black")
```

## pairwise among depth bands

```{r}
logm_sarg96_mod |> 
emmeans(~ Depth_band,
        type = "link") |> 
  pairs() |> 
  gather_emmeans_draws() |> # odds-ratio
  mutate(.value = exp(.value)) |> # log odds-ratio  
  summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1)) |> 
  filter(Pl > 0.85 | Pg > 0.85)
```

### ggsave macroalgal assay

```{r}
ggsave("Plots/Fig. X. sarg_assays_96h.jpeg",
       plot = p_sarg_assay_96, 
       dpi = 600, 
       height = 4, 
       width = 6, 
       units = "in")
```

# 3H EXPERIMENTS

## load model

```{r}
load("Model_Outputs/logm_sarg3h_mod.Rdata")
```

## summary to check site variation

```{r}
summary(logm_sarg3h_mod)

# response scale of site variation
exp(0.23)
```

## summary table

```{r}
logm_sarg3h_mod |> as_draws_df() |> 
  dplyr::select(matches("^b_.*")) |> 
  exp() |> 
  summarise_draws(median,
                  HDInterval::hdi,
                  Pl= ~mean(.x < 1),
                  Pg = ~mean(.x > 1)) #|> 
  #filter(Pl > 0.85 | 
   #        Pg > 0.85)
```

## newdata for plotting macroalgal loss with depth bands

```{r}
sarg_assay_3h_nd <-
  logm_sarg3h_mod |> 
  emmeans(~ Depth_band *
            Treatment, 
          type = "link") |> 
  gather_emmeans_draws() |> 
  mutate(Exp_grams = exp(.value)) |> 
  dplyr::select(-.chain,
                -.iteration,
                -.value) |> 
  summarise_draws(median, 
                  HDInterval::hdi) |> 
  cbind(vid_replicates = c(12, 12, 12, 12, 12, 12, 6, 6)) |> 
  as.data.frame() |> 
  mutate(Depth_band = recode(Depth_band,
                             "10m" = "0-10m",
                             "20m" = "11-20m",
                             "30m" = "30-40m",
                             "50m" = "41-50m"))
```

# check values for open treatments

```{r}
sarg_assay_3h_nd |> filter(Treatment == "Open")
```

## plot

```{r}
p_sarg_assay_3h <- 
sarg_assay_3h_nd |> 
  ggplot(aes(x = Depth_band,
             y = median,
             fill = Treatment)) +
  geom_pointrange(aes(ymin = lower, 
                      ymax = upper,
                      shape = Treatment),
                  position = position_dodge(0.30),
                  size = 1.5,
                  fatten = 3) +
  scale_y_continuous(expression(italic(Sargassum)~loss~(grams~3~h^-1)),
                     limits = c(0,15),
                     breaks = seq(0, 15, by = 5)) +
  scale_x_discrete(name = "Depth band") +
  theme_classic() +
  ggtitle("Macroalgal video assays") +
  scale_fill_manual(values = c("red2", "blue")) +
   scale_shape_manual(values = c(22, 21)) +
  theme(axis.title.y = element_text(size = 15),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(size = 18),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 15)) +
  geom_image(data = tibble(Depth_band = 0.90,
                           median = 12,
                           Treatment = "Caged"),
             aes(image = "Images/sargassum-spp.png"),
             size = 0.30) +
   geom_text(aes(label = vid_replicates,
                  y = upper + 0.5,
                  x = ),
              position = position_dodge(width = dodge_width),
              color = "black")
```

## pairwise among depth bands

```{r}
logm_sarg3h_mod |> 
emmeans(~ Depth_band,
        type = "link") |> 
  pairs() |> 
  gather_emmeans_draws() |> # odds-ratio
  mutate(.value = exp(.value)) |> # log odds-ratio  
  summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1)) |> 
  filter(Pl > 0.85 | Pg > 0.85)
```

### end
