---
title: "bites_results_plots"
format: html
editor: visual
---

# Description

This qmd file examines the models from the "fits" qmd, which investigated changes in bite rates for two Naso species among depth bands

# NASO UNICORNIS

## load model

```{r}
load("Model_Output/nu_mod_hurdle.Rdata")
```

## summary

```{r}
 nu_mod_hurdle |>  
  as_draws_df() |> 
  dplyr::select(matches("^b_.*")) |>
  exp() |> 
  summarise_draws(median,
                  HDInterval::hdi,
                  Pl= ~mean(.x < 1),
                  Pg = ~mean(.x > 1))
```

## newdata for NU plotting

```{r}
nu_nd <- 
  nu_mod_hurdle |> 
  emmeans(~ Depth_band,
          type = "link") |> 
  gather_emmeans_draws() |> 
  mutate(Exp_bites = exp(.value)) |> 
  dplyr::select(-.chain,
                -.iteration,
                -.value) |> 
  summarise_draws(median, 
                  HDInterval::hdi) |> 
    cbind(vid_replicates = c(12, 12, 12, 6)) |> 
  as.data.frame()
```

## correct values for Nu: 0 for 20m since no observation at 20m

```{r}
nu_nd <- 
nu_nd |> 
  rows_delete(tibble(Depth_band = "20m")) |> 
  add_row(Depth_band = "20m",
           variable = "Exp_bites", 
           median = 0,
           lower = 0,
           upper = 0,
          vid_replicates = 12) |> 
  mutate(
         Depth_band = factor(Depth_band,
                             levels = c("10m",
                                        "20m",
                                        "30m",
                                        "50m")))
```

## pairwise comparisons among depth bands

```{r}
 nu_mod_hurdle |> 
emmeans(~ Depth_band,
        type = "link") |>
  pairs() |> 
  gather_emmeans_draws() |> # odds-ratio
  mutate(.value = exp(.value)) |> # log odds-ratio  
  summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1))
```

# plot Nu

```{r}
p_bites_nu <-
nu_nd |> 
  mutate(Depth_band = recode(Depth_band,
                             "10m" = "0-10m",
                             "20m" = "11-20m",
                             "30m" = "30-40m",
                             "50m" = "41-50m")) |> 
  ggplot(aes(x = Depth_band,
             y = median)) +
  geom_pointrange(aes(ymin = lower,
                      ymax = upper),
                  shape = 21,
                  size = 1.5,
                  fatten = 3,
                  color = "black",
                  fill = "red2") +
  scale_y_continuous(expression(Number~of~bites~(3~h^-1)),
                     limits = c(0, 250),
                     breaks = seq(0, 250, by = 50)) +
  scale_x_discrete(name = "Depth bands") +
  theme_classic() +
    ggtitle(expression(italic(Naso~unicornis))) +
     theme(
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 12),
        plot.title = element_text(size = 18)) +
  geom_image(data = tibble(Depth_band = 1,
                           median = 235),
             aes(image = "Images/naso-unicornis-bluespine-unicornfish.png"),
             size = 0.40) +
  geom_text(aes(label = vid_replicates,
                  y = upper + 15,
                  x = ),
              position = position_dodge(width = dodge_width),
              color = "black")
```

# NASO LITURATUS

## load model

```{r}
load("Model_Output/nl_mod_hurdle.Rdata")
```

# summary table

```{r}
 nl_mod_hurdle |>  
  as_draws_df() |> 
  dplyr::select(matches("^b_.*")) |>
  exp() |> 
  summarise_draws(median,
                  HDInterval::hdi,
                  Pl= ~mean(.x < 1),
                  Pg = ~mean(.x > 1))
```

# pairwise comparisons among depth bands

```{r}
 nl_mod_hurdle |> 
emmeans(~ Depth_band,
        type = "link") |>
  pairs() |> 
  gather_emmeans_draws() |> # odds-ratio
  mutate(.value = exp(.value)) |> # log odds-ratio  
  summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1))
```

# newdata for Nl plotting

```{r}
nl_nd <- 
  nl_mod_hurdle |> 
  emmeans(~ Depth_band,
          type = "link") |> 
  gather_emmeans_draws() |> 
  mutate(Exp_bites = exp(.value)) |> 
  dplyr::select(-.chain,
                -.iteration,
                -.value) |> 
  summarise_draws(median, 
                  HDInterval::hdi) |> 
    cbind(vid_replicates = c(12, 12, 12, 6)) |> 
  as.data.frame()
```

# correct values for Nl: 0 for 30m since no observation at 30m

```{r}
nl_nd <- 
nl_nd |> 
  rows_delete(tibble(Depth_band = "30m")) |> 
  add_row(Depth_band = "30m",
           variable = "Exp_bites", 
           median = 0,
           lower = 0,
           upper = 0,
          vid_replicates = 12) |> 
  mutate(
         Depth_band = factor(Depth_band,
                             levels = c("10m",
                                        "20m",
                                        "30m",
                                        "50m")))
```

# plot Nl

```{r}
p_bites_nl <-
nl_nd |> 
  mutate(Depth_band = recode(Depth_band,
                             "10m" = "0-10m",
                             "20m" = "11-20m",
                             "30m" = "30-40m",
                             "50m" = "41-50m")) |> 
  ggplot(aes(x = Depth_band,
             y = median)) +
  geom_pointrange(aes(ymin = lower,
                      ymax = upper),
                  shape = 21,
                  size = 1.5,
                  fatten = 3,
                  color = "black",
                  fill = "blue") +
  scale_y_continuous(expression(Number~of~bites~(3~h^-1)),
                     limits = c(0, 250),
                     breaks = seq(0, 250, by = 50)) +
  #scale_x_discrete(name = "") +
  theme_classic() +
    ggtitle(expression(italic(Naso~lituratus))) +
     theme(axis.title.x = element_blank(),
           axis.text.x = element_blank(),
           axis.ticks.x = element_blank(),
        axis.title.y = element_text(size = 15),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 18)) +
  geom_image(data = tibble(Depth_band = 1,
                           median = 240),
             aes(image = "Images/naso-lituratus-naso-tang.png"),
             size = 0.40) +
   geom_text(aes(label = vid_replicates,
                  y = upper + 14,
                  x = ),
              position = position_dodge(width = dodge_width),
              color = "black")
```

# COMBINE plots

```{r}
p_vid_assays <- 
  p_sarg_assay_3h /
  p_bites_nl /
  p_bites_nu + plot_annotation(tag_levels = list(c("A)",
                                                   "B)",
                                                   "C)"))) &
  theme(plot.tag = element_text(size = 15))
```

## ggsave

```{r}
ggsave("Plots/Fig. X. bites.jpeg",
       plot = p_vid_assays, 
       dpi = 600, 
       height = 12, 
       width = 10, 
       units = "in")
```

# FOR PRECOMPLETION

# plot Nu

```{r}
p_bites_nu <-
nu_nd |> 
  mutate(Depth_band = recode(Depth_band,
                             "10m" = "0-10m",
                             "20m" = "11-20m",
                             "30m" = "30-40m",
                             "50m" = "41-50m")) |> 
  ggplot(aes(x = Depth_band,
             y = median)) +
  geom_pointrange(aes(ymin = lower,
                      ymax = upper),
                  shape = 21,
                  size = 3,
                  fatten = 3,
                  color = "black",
                  fill = "red2") +
  scale_y_continuous(expression(Number~of~bites~(3~h^-1)),
                     limits = c(0, 75),
                     breaks = seq(0, 75, by = 25)) +
  scale_x_discrete(name = "Depth bands") +
  theme_classic() +
    ggtitle(expression(italic(Naso~unicornis))) +
     theme(
        axis.title = element_text(size = 30),
        axis.text = element_text(size = 28),
        plot.title = element_text(size = 32)) +
  geom_image(data = tibble(Depth_band = 1,
                           median = 70),
             aes(image = "Images/naso-unicornis-bluespine-unicornfish.png"),
             size = 0.40)
```

## ggsave

```{r}
ggsave("Plots/p_bites_nu.jpeg",
       plot = p_bites_nu, 
       dpi = 600, 
       height = 8, 
       width = 12, 
       units = "in")
```

# plot Nl

```{r}
p_bites_nl <-
nl_nd |> 
  mutate(Depth_band = recode(Depth_band,
                             "10m" = "0-10m",
                             "20m" = "11-20m",
                             "30m" = "30-40m",
                             "50m" = "41-50m")) |> 
  ggplot(aes(x = Depth_band,
             y = median)) +
  geom_pointrange(aes(ymin = lower,
                      ymax = upper),
                  shape = 21,
                  size = 3,
                  fatten = 3,
                  color = "black",
                  fill = "blue") +
  scale_y_continuous(expression(Number~of~bites~(3~h^-1)),
                     limits = c(0, 250),
                     breaks = seq(0, 250, by = 50)) +
  scale_x_discrete(name = "Depth band") +
  theme_classic() +
    ggtitle(expression(italic(Naso~lituratus))) +
     theme(
        axis.title = element_text(size = 30),
        axis.text = element_text(size = 28),
        plot.title = element_text(size = 32)) +
  geom_image(data = tibble(Depth_band = 1,
                           median = 240),
             aes(image = "Images/naso-lituratus-naso-tang.png"),
             size = 0.40)
```

## ggsave

```{r}
ggsave("Plots/p_bites_nl.jpeg",
       plot = p_bites_nl, 
       dpi = 600, 
       height = 8, 
       width = 12, 
       units = "in")
```
