---
title: "benthic_cover_results_plots"
format: html
editor: visual
---

# Description

This qmd takes the saved model outputs from "fits" qmd and investigates the evidence for an effect of depth and visualizes the output in ggplot.

# CORAL

# load model

```{r}
load("Model_Output/coral_mod_hab.Rdata")
```

## summary table to detect evidence for an effect

```{r}
coral_mod |> as_draws_df() |> 
  dplyr::select(matches("^b_.*")) |>
  exp() |> 
  summarise_draws(median,
                  HDInterval::hdi,
                  Pl= ~mean(.x < 1),
                  Pg = ~mean(.x > 1))
```

## check coral values for depth band with emmeans

```{r}
coral_mod |> 
emmeans(~ Depth_band,
        type = "link") |> 
  gather_emmeans_draws() |> 
  mutate(prob = plogis(.value),
         percentage = prob * 100) |> 
  dplyr::select(-.chain,
                -.iteration,
                -.draw,
                -.value) |> 
  summarise_draws(median,
                  HDInterval::hdi,
                  Pl = ~mean(.x < 1),
                  Pg = ~mean(.x > 1)) |> 
  filter(variable == "percentage")
```

## plot coral across depth bands

```{r}
p_coral <- 
coral_mod |> 
emmeans(~ Depth_band,
        type = "link") |> 
  gather_emmeans_draws() |> 
  mutate(prob = plogis(.value),
         percentage = prob * 100) |> 
  dplyr::select(-.chain,
                -.iteration,
                -.draw,
                -.value) |> 
  summarise_draws(median,
                  HDInterval::hdi,
                  Pl = ~mean(.x < 1),
                  Pg = ~mean(.x > 1)) |> 
  filter(variable == "percentage") |> 
  cbind(bc_replicates = c(12, 12, 12, 6)) |>
  ggplot(aes(x = Depth_band,
             y = median)) +
  geom_pointrange(aes(ymin = lower, 
                      ymax = upper),
                  position = position_dodge(0.25),
                  shape = 21,
                  size = 1.5,
                  fatten = 3,
                  color = "black",
                  fill = "red2") +
  scale_y_continuous(limits = c(0,80),
                     breaks = seq(0, 80, by = 20),
                     expression(Percentage~cover)) +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        plot.title = element_text(size = 18),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 15)) +
  theme(legend.position = "none") +
  ggtitle("Coral") +
  geom_image(data = tibble(Depth_band = 4.2,
                           median = 74),
             aes(image = "Images/acropora-spp.png"),
             size = 0.30) +
   geom_text(aes(label = bc_replicates,
                  y = upper + 3,
                  x = ),
              position = position_dodge(width = dodge_width),
              color = "black")
```

## pairwise table to compare among depth bands

```{r}
#coral_mod_pairs <- 
coral_mod |> 
emmeans(~ Depth_band,
        type = "link") |>
  pairs() |> 
  gather_emmeans_draws() |> # log-odds ratio
  mutate(.value = exp(.value)) |>    
  summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1))#,
            #.width = c(0.85, 0.95))
```

# MACROALGAE

# load model

```{r}
load("Model_Output/ma_mod_hab.Rdata")
```

## summary table

```{r}
ma_mod |> as_draws_df() |> 
  dplyr::select(matches("^b_.*")) |>
  exp() |> 
  summarise_draws(median,
                  HDInterval::hdi,
                  Pl= ~mean(.x < 1),
                  Pg = ~mean(.x > 1))
```

## check macroalgae values for depth band with emmeans

```{r}
ma_mod |> 
emmeans(~ Depth_band,
        type = "link") |> 
  gather_emmeans_draws() |> 
  mutate(prob = plogis(.value),
         percentage = prob * 100) |> 
  dplyr::select(-.chain,
                -.iteration,
                -.draw,
                -.value) |> 
  summarise_draws(median,
                  HDInterval::hdi,
                  Pl = ~mean(.x < 1),
                  Pg = ~mean(.x > 1)) |> 
  filter(variable == "percentage")
```

## plot macroalgae across depth bands

```{r}
p_macroalgae <- 
ma_mod |> 
emmeans(~ Depth_band,
        type = "link") |> 
  gather_emmeans_draws() |> 
  mutate(prob = plogis(.value),
         percentage = prob * 100) |> 
  dplyr::select(-.chain,
                -.iteration,
                -.draw,
                -.value) |> 
  summarise_draws(median,
                  HDInterval::hdi,
                  Pl = ~mean(.x < 1),
                  Pg = ~mean(.x > 1)) |> 
  filter(variable == "percentage") |> 
  cbind(bc_replicates = c(12, 12, 12, 6)) |> 
  mutate(Depth_band = recode(Depth_band,
                             "10m" = "0-10m",
                             "20m" = "11-20m",
                             "30m" = "31-40m",
                             "50m" = "41-50m")) |> 
  ggplot(aes(x = Depth_band,
             y = median)) +
  geom_pointrange(aes(ymin = lower, 
                      ymax = upper),
                  position = position_dodge(0.25),
                  shape = 21,
                  size = 1.5,
                  fatten = 3,
                  color = "black",
                  fill = "blue") +
  scale_y_continuous(limits = c(0,80),
                     breaks = seq(0, 80, by = 20),
                     expression(Percentage~cover)) +
  theme_classic() +
  theme(
        plot.title = element_text(size = 18),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 15)) +
  scale_x_discrete(name = "Depth bands") +
  theme(legend.position = "none") +
  ggtitle("Macroalgae") +
  geom_image(data = tibble(Depth_band = 4.2,
                           median = 64),
             aes(image = "Images/sargassum-spp.png"),
             size = 0.30) +
   geom_text(aes(label = bc_replicates,
                  y = upper + 3,
                  x = ),
              position = position_dodge(width = dodge_width),
              color = "black")
```

## pairwise table to compare among depth bands

```{r}
#coral_mod_pairs <- 
ma_mod |> 
emmeans(~ Depth_band,
        type = "link") |>
  pairs() |> 
  gather_emmeans_draws() |> # odds-ratio
  mutate(.value = exp(.value)) |> # log odds-ratio  
  summarise(median_hdci(.value),
            Pl = mean(.value < 1),
            Pg = mean(.value > 1))#,
            #.width = c(0.85, 0.95))
```

# COMBINE: benthic plots

```{r}
p_benthic <-
p_coral /
p_macroalgae +
  plot_annotation(tag_levels = list(c("A)", "B)"))) &
  theme(plot.tag = element_text(size = 15))
```

## ggsave

```{r}
ggsave("Plots/Fig. X. coral_macroalgae.jpeg",
       plot =p_benthic, 
       dpi = 600, 
       height = 8, 
       width = 8, 
       units = "in")
```

### end
